---
title: "cDNA_Ochre_analysis"
author: "MT"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ORGANIZE YOUR PROJECT FOLDERS
You should keep the organization of each project folder the same. This will ensure that scripts are reproducible across different projects (where you pull data from, where you write objects or plots to etc.)

Files to use are:
- Data
- Output
  - Plots (sub folder)
- Scripts


#Housekeeping

```{r}
date()
Sys.time()
getwd()
list.files()
.libPaths()
.Library
sessionInfo()
search()
searchpaths()
R.version


```

# Checking and installing packages

```{r, results = 'hide'}

packages <- c("ggplot2", "dplyr", "tidyr", "readxl", "ggpubr", "skimr", "DataExplorer", "tidyverse", "skimr", "svglite", "readxl","tidyxl", "ggforce", "ggpubr", "ggsci", "ggthemes", "ragg", "magrittr", "reprex", "patchwork", "writexl", 'magick', 'forcats')

# list of packages sometimes I use [   "searcher", "errorist",

install.packages(setdiff(packages, rownames(installed.packages())))

update.packages (packages)

for (i in 1:length(packages)) {

library(packages[i], character.only = TRUE)


}

```

```{r}
# Import data and perform initial cleaning
dataFinal <- read_csv("2023_Data/Data/20230621FISH_LZ22221.csv") %>%
  mutate(strain = as.factor(strain),
         cond = as.factor(cond),
         avgInt = totInt / cellArea,
         id = paste(cond, "_", strain, "_", time, sep = "")) %>%
  select(totInt, cellArea, avgInt, cellnum, bkgrdInt, frame, cond, time, strain, id) %>%
  mutate(id = as.factor(id))


# Plot histograms for numerical columns
purrr::walk(names(select_if(dataFinal, is.numeric)), ~{
  hist(dataFinal[[.]], main = paste("Histogram of", .), xlab = "Value", col = "lightblue", breaks = 40)
})


# Plot histogram of average intensity, faceted by id
ggplot(data = dataFinal, mapping = aes(x = avgInt)) + 
  geom_histogram() + 
  facet_wrap(~id)

# # Filter data for condition
# strainTest <- dataFinal %>%
#   filter(cond == 'ochre') %>%
#   summarise(meanBkgd = mean(unique(bkgrdInt)),
#             sdBkgd = sd(unique(bkgrdInt)))

# Plot histogram of average intensity (corrected for background), faceted by time
ggplot(data = dataFinal %>%
         filter(avgInt > 2734), 
       mapping = aes(x = avgInt - 2734, fill = cond)) + 
  geom_histogram(binwidth = 700) + 
  facet_wrap(~time)

# Calculate and print the percentage of fluorescent cells for each time point
dataFinal %>%
  group_by(cond,time) %>%
  summarise(totalCells = n(),
            fluorescentCells = sum(avgInt > 2822.568),
            percentageFluorescent = 100 * fluorescentCells / totalCells) %>%
  rowwise() %>%
  mutate(result = paste("For time ", time, 
                        " total number of cells: ", totalCells, 
                        " Number of fluorescent cells: ", fluorescentCells, 
                        " Percent of cells that fluoresce: ", percentageFluorescent)) %>%
  pull(result) %>%
  walk(print)



# Summary of data by condition and strain
summary_table <- dataFinal %>%
  group_by(cond, strain, time) %>%
  summarise(
    n = n(),
    avg_totInt = mean(totInt),
    avg_cellArea = mean(cellArea),
    avg_avgInt = mean(avgInt),
    sd_totInt = sd(totInt),
    sd_cellArea = sd(cellArea),
    sd_avgInt = sd(avgInt),
    avg_bkgrd = mean(bkgrdInt),
    sd_bkgrd = sd(bkgrdInt),
    .groups = 'drop'
  )

print(summary_table)
```

From the summary_table we see that our cell numebr for each time point/condition is highly variable which could lead to a skew in teh data analysis allowing for smaller populations of cells to "pull" our values into different directions. We may be able to account for this in our analysis but it is always easiest to first look at data collection and make sure the data collection is being done in a way without biasing our data and in a way that allows for the most streamlines analysis (least variability and need for cleaning)




```{r}

# Print summary table
print(summary_table)

# Boxplot of average intensity by condition
ggplot(dataFinal, aes(x = cond, y = avgInt)) +
  geom_boxplot() +
  labs(title = "Boxplot of Average Intensity by Condition",
       x = "Condition",
       y = "Average Intensity")

# Correlation between total intensity and cell area
correlation <- cor.test(dataFinal$totInt, dataFinal$cellArea)

# Print correlation
print(paste("Correlation between total intensity and cell area: ", correlation$estimate))

```



```{r}
# Additional Data Analysis






```
